Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-06-03 talk/dls2014/paper/paper.tex
clean up footnotes
the full stop goes before the footnote

\lstinline!stm_start_transaction()!  starts a transaction. It requires
two arguments, the first being a thread-local data structure and the
second a buffer for use by \lstinline!setjmp()!.
\lstinline!stm_commit_transaction()!  tries to commit the current
transaction. \lstinline!stm_read()!, \lstinline!stm_write()!  perform
a read or a write barrier on an object and \lstinline!stm_allocate()!
allocates a new object with the specified size.
 \lstinline!STM_PUSH_ROOT()!  and \lstinline!STM_POP_ROOT()!  push
and pop objects on the shadow stack~\footnote{A stack for pointers to
  GC objects that allows for precise garbage collection. All objects
  on that stack are never seen as garbage and are thus always kept
  alive.~\cite{fergus02}}.  Objects have to be saved using this stack around calls
that may cause a GC cycle to happen, and also while there is no
transaction running. In this simplified API, only
\lstinline!stm_allocate()!  and \lstinline!stm_commit_transaction()!
require saving object references.

The type \lstinline!object_t!  is special as it causes the
compiler~\footnote{Clang 3.5 with some patches to its address-space
 256 feature} to make all accesses through it relative to the $\%gs$
register.  With exceptions, nearly all accesses to objects managed by
the TM system should use this type so that the CPU will translate the
reference to the right version of the object.

@@

In all cases, we remove the \lstinline!WRITE_BARRIER!  flag from the
object before we return. Thus, we never trigger the slow path again
before we do the next minor collection or we start the next
transaction~\footnote{We always do a minor collection during a commit.}.

@@

In Figure~\ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards\footnote{OS kernel simulation
benchmark}. The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second.

>>>>

\lstinline!stm_start_transaction()!  starts a transaction. It requires
two arguments, the first being a thread-local data structure and the
second a buffer for use by \lstinline!setjmp()!.
\lstinline!stm_commit_transaction()!  tries to commit the current
transaction. \lstinline!stm_read()!, \lstinline!stm_write()!  perform
a read or a write barrier on an object and \lstinline!stm_allocate()!
allocates a new object with the specified size.
 \lstinline!STM_PUSH_ROOT()!  and \lstinline!STM_POP_ROOT()!  push
and pop objects on the shadow stack.\footnote{A stack for pointers to
  GC objects that allows for precise garbage collection. All objects
  on that stack are never seen as garbage and are thus always kept
  alive.~\cite{fergus02}}  Objects have to be saved using this stack around calls
that may cause a GC cycle to happen, and also while there is no
transaction running. In this simplified API, only
\lstinline!stm_allocate()!  and \lstinline!stm_commit_transaction()!
require saving object references.

The type \lstinline!object_t!  is special as it causes the
compiler\footnote{Clang 3.5 with some patches to fix bugs in its address-space
 256 feature} to make all accesses through it relative to the $\%gs$
register.  With exceptions, nearly all accesses to objects managed by
the TM system should use this type so that the CPU will translate the
reference to the right version of the object.

@@

In all cases, we remove the \lstinline!WRITE_BARRIER!  flag from the
object before we return. Thus, we never trigger the slow path again
before we do the next minor collection or we start the next
transaction.\footnote{We always do a minor collection during a commit.}

@@

In Figure~\ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards.\footnote{Richards is a OS kernel simulation
benchmark.} The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second.
